---
title: "Lab 06 - Sad plots"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```
### Prepare for Exercise

```{r}
staff <- read_csv("data/instructional-staff.csv")
print(staff)


```

Data reshaping 
```{r}
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value)) %>%
  print()

```

```{r}
staff_long %>%
 
  ggplot(aes(x = year, y = value, fill = faculty_type)) +
  theme(axis.text.x=element_text(angle=60)) +
  geom_col(position="stack")

```
### Exercise 1

```{r}
staff_long$faculty_type<-as.factor(staff_long$faculty_type)
staff_long$faculty_type <- relevel(staff_long$faculty_type,"Full-Time Tenured Faculty","Full-Time Tenure-Track Faculty", "Full-Time Non-Tenure-Track Faculty", "Part-Time Faculty","Graduate Student Employees")

```

```{r,fig.width=8}

staff_long %>%
  ggplot(aes(x = year, y = value, group = faculty_type, color = faculty_type)) +
  geom_line() +
  ylim(0, max(staff_long$value)) +
  labs (title="Faculty type and its changes across the time",
        x= "Year", y="Ratio", color="type of faculty")+
  theme(axis.text.x=element_text(angle=60))+
  scale_colour_brewer(palette="Set1")



```



### Exercise 2

```{r,fig.width=6}
staff_long %>%
  ggplot(aes(x = year, y = value, fill = faculty_type)) +
  geom_col(position="dodge")+
  theme(axis.text.x=element_text(angle=60))+
  scale_fill_manual(
    values = c("Full-Time Tenured Faculty" = "gray10", "Full-Time Non-Tenure-Track Faculty" = "gray20", "Full-Time Tenure-Track Faculty"="gray30","Graduate Student Employees"="gray40", "Part-Time Faculty"="deeppink2")) +
  labs (title="Faculty type and its changes across the time",
        x= "Year", y="Ratio", fill="type of faculty")


```
### Exercise 3
```{r}
fisheries <- read_csv("data/fisheries.csv")

fisheries <- fisheries %>%
  dplyr::filter(total>100000)

```


```{r}

library(maps)

library(mapdata)

library(rworldmap)




fish_cap<- joinCountryData2Map(fisheries, joinCode="NAME", nameJoinColumn="country")

cap1<-mapCountryData(fish_cap, nameColumnToPlot="capture", catMethod="fixedWidth", addLegend = TRUE, missingCountryCol="black")


cap2<-mapCountryData(fish_cap, nameColumnToPlot="aquaculture", catMethod="fixedWidth",addLegend = TRUE, missingCountryCol="black")



```


```{r,fig.width=10}
# tried messy graph but better to erase it  

# fisheries_lon <- fisheries %>%
#   pivot_longer(cols=-c("country","total"), names_to="type")%>%
#   mutate(value=as.numeric(value))%>%
#   print(head(100))
# 

```
## Stretch Practice with Smokers in Whickham


```{r}
#install.packages('mosaicData')
library(tidyverse) 
library(mosaicData) 

```

```{r}

data(Whickham)
Whickham<-Whickham
#?Whickham
```

1. observation. 
because we cannot make people smoke for a life, and we cannot manipulate their smoke habit, age, death status. 
2. 1314, # of participants. 
3. 3
4. I think smokers would die early :( 


```{r}

Whickham %>%
  count(smoker, outcome)
```

5. I think it does not make a cogent case to draw the conclusion on smoking-health relationship from the above table, because I think Age is really important variable for death status (it is not health status I think). 

```{r}
  
Whickham%>%
  ggplot(aes(x=outcome, fill=smoker)) +
  geom_bar( position="dodge")

```

```{r}

data(Whickham)
?Whickham
```
### Exercise 5
```{r}

  
Whickham%>%
  ggplot(aes(x=outcome, fill=smoker)) +
  geom_bar( position="dodge")

## This graph does not give any explanations on possibilities. 
## it only depicts the numbers. 
## So I tried ratio-manner approaches. 

 Whickham2 <-Whickham %>%
  count(smoker, outcome) %>%
  group_by(outcome) %>%
  dplyr::mutate(prob = n/sum(n))
 
 Whickham2%>%
  ggplot(aes(x=outcome, y=prob, fill=smoker)) +
  geom_bar(stat="identity")


```
### Exercise 6
```{r}

Whickham <-Whickham %>%
  dplyr::mutate(age_cat = case_when(
    age <= 44 ~ "18-44",
    age > 44 & age <= 64 ~ "45-64",
    age > 64 ~ "65+",
    TRUE ~ NA_character_))

Whickham3 <-Whickham %>%
  count(smoker, age_cat, outcome)%>%
  group_by(outcome) %>%
  dplyr::mutate(prob = n/sum(n))

Whickham3 %>%
  ggplot(aes(x=outcome, y=prob, fill=smoker)) +
  facet_grid(~age_cat)+
  geom_bar(stat="identity")

Whickham3 %>%
  ggplot(aes(x=outcome,  fill=smoker, y=n)) +
  facet_grid(~age_cat)+
  geom_histogram(stat="identity", position="dodge")

  
```
